---
title: "AE 03: Duke Forest + data visualization"
author: Zachary Kujala
editor: 
  markdown: 
    wrap: sentence
---

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
```

## Exercise 1

Suppose you're helping some family friends who are looking to buy a house in Duke Forest.
As they browse Zillow listings, they realize some houses have garages and others don't, and they wonder: **Does having a garage make a difference?**

Luckily, you can help them answer this question with data visualization!

-   Make histograms of the prices of houses in Duke Forest based on whether they have a garage.
    -   In order to do this, you will first need to create a new variable called `garage` (with levels `"Garage"` and `"No garage"`).
    -   Below is the code for creating this new variable. Here, we `mutate()` the `duke_forest` data frame to add a new variable called `garage` which takes the value `"Garage"` if the text string `"Garage"` is detected in the `parking` variable and takes the test string `"No garage"` if not.

```{r}
#| label: garage
#| eval: false

duke_forest |>
  mutate(garage = if_else(str_detect(parking, "Garage"),   "Garage", "No garage"))
```

-   Then, facet by `garage` and use different colors for the two facets.
-   Choose an appropriate binwidth and decide whether a legend is needed, and turn it off if not.
-   Include informative title and axis labels.
-   Finally, include a brief (2-3 sentence) narrative comparing the distributions of prices of Duke Forest houses that do and don't have garages. Your narrative should touch on whether having a garage "makes a difference" in terms of the price of the house.

```{r}
#| label: price-garage

duke_forest <- duke_forest |> 
  mutate(garage = if_else(str_detect(parking, "Garage"), "Garage", "No garage"))

ggplot(duke_forest, aes(x = price, fill = garage)) +
  facet_wrap(~garage) +
  geom_histogram(binwidth = 80000) +
  labs(
    x = "Price ($)",
    y = "Number of Houses",
    title = "Relationship between Garages and House Prices"
  ) + 
  theme(legend.position = "none")
```

The addition of a garage does not appear to have a significant impact on house price.
The mode price of houses with and without garages are both approximately \$550,000.
In fact, the highest priced house is one without a garage.
Since both distributions are similar we cannot say if having a garage makes a difference on the price of houses.

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::

## Exercise 2

It's expected that within any given market larger houses will be priced higher.
It's also expected that the age of the house will have an effect on the price.
However in some markets new houses might be more expensive while in others new construction might mean "no character" and hence be less expensive.
So your family friends ask: "In Duke Forest, do houses that are bigger and more expensive tend to be newer ones than those that are smaller and cheaper?"

Once again, data visualization skills to the rescue!

-   Create a scatter plot to exploring the relationship between `price` and `area`, conditioning for `year_built`.
-   Use `geom_smooth()` with the argument `se = FALSE` to add a smooth curve fit to the data and color the points by `year_built`.
-   Include informative title, axis, and legend labels.
-   Discuss each of the following claims (1-2 sentences per claim). Your discussion should touch on specific things you observe in your plot as evidence for or against the claims.
    -   Claim 1: Larger houses are priced higher.
    -   Claim 2: Newer houses are priced higher.
    -   Claim 3: Bigger and more expensive houses tend to be newer ones than smaller and cheaper ones.

```{r}
#| label: price-area
#| message: false
#| warning: false

ggplot(duke_forest, aes(x = price, y = area, color = year_built)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(
    x = "Price ($)",
    y = "Area",
    title = "House Prices and Areas by Year",
    color = "Year Built"
  )
```

**Claim 1:** Based on the plot some houses with larger areas are higher priced, seen by the top right data points and the positive curve.
However there are many houses that have smaller areas yet are priced the same as ones with larger areas, as seen by the cluster of data points between the \$400,000 to \$800,000 price range.
Also there are houses with larger areas that are cheap, seen by the highest data points in the \$0 to \$600,000 price range.

**Claim 2:** Newer houses are not priced higher.
Besides the outlier in the top right corner of the plot, all the newest houses have a price of about \$600,000 or less.
The houses above this price are all older.

**Claim 3:** Bigger and more expensive houses do not tend to be newer ones.
The exception to this is the one outlier in the top right corner of the plot which does support the claim.
The plot displays that newer houses are typically under \$600,000 and have smaller areas than many older houses.

::: callout-important
Now is a good time to render, commit, and push.
Make sure that you commit and push all changed documents and your Git pane is completely empty before proceeding.
:::
